Tidy data structure to support exploration and modeling of temporal-context data

### 2018-11-06 main.Rmd
added new castle smart meter data 2013

## Time index and interval {#sec:index}

@@

@@

The "key" identifies units or subjects that are observed over time in a data table, which is typically a priori known structure by users. The absence of key only occurs to a univariate case where the key is implicit. Multiple units must associate with an explicit "key" to allow for identification. The key is not constrained to a single column, but is comprised of multiple columns. When involving multiple variables, one need to distinguish nesting and crossing data variables. "Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`). Nesting is a special case of crossing. The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference. Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data. Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics. The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables. @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions. A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout. However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.

@@

The ground of time series modelling or forecasting is left untouched in this paper. The future plan is to bridge the gap between tidy data and model building. Currently, it is required to casting to matrix from tidy data and therefore building a model. But time series models should be directly applied to tidy data as other wrangling tools do, without such an intermediate step. In particular, a univariate time series model, like ARIMA and Exponential Smoothing, can be applied to multiple time series independently. A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future. Model summaries include coefficients, fitted values, and residuals; forecasting objects include future time path and distributions generating prediction intervals.

>>>>

## Index {#sec:index}

@@

The "key" specification contributes to forming a tsibble alongside the index. The concept of "key" is introduced to identify units or subjects that are observed over time in a data table. In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification. In the long format, all column names are melt to the corresponding "key" values. But the "key" is much more flexible than simply column names. Because it is not constrained to a single field, but can be comprised of multiple fields. The variables remain the same as it is, without additional manipulation to create the identifiers.

It is typically a priori known by users. The absence of key only occurs to a univariate case where the key is implicit. Multiple units must associate with an explicit "key" to allow for identification.

## Interval {#sec:interval}

@@

@@

The ground of time series modelling or forecasting is left untouched in this paper. The future plan is to bridge the gap between tidy data and model building. Currently, it is required to casting to matrix from tidy data and therefore building a model. But time series models should be directly applied to tidy data as other wrangling tools do, without such an intermediate step. In particular, a univariate time series model, like ARIMA and Exponential Smoothing, can be applied to multiple time series independently. A tidy format to represent model summaries and forecasting objects will be developed and implemented in the future. Model summaries include coefficients, fitted values, and residuals; forecasting objects.

When involving multiple variables, one need to distinguish nesting and crossing data variables. "Nesting" refers to a scenario where a variable is nested within another when each category of the former variable co-occurs with only one category of the latter (for example, `country` nested within `continent`); on the other hand, "crossing" means that a variable is crossed with another when every category of one variable co-occurs with every category of the other (for example, `country` crosses with `gender`). Nesting is a special case of crossing. The former exhibits a hierarchical ordering with lower-level category (children) nested within higher-level category (parent) (more categories in lower level); whereas the ordering of the latter makes no difference. Whilst constructing tidy temporal data, users also need to specify the key that identifies the structure on the data. Rather than creating a separate object that contains the data structure, like what the **hts** does, we propose a syntax-based interface to directly deal with variables, so that it takes advantage of data semantics. The expression of a vertical bar (`|`) (can be interpreted probabilistically as the conditional distribution for its parent) is used to indicate nesting variables (reflect depths), while a comma (`,`) for crossing variables. @wilkinson2006grammar discussed the distinction between nesting and crossing with respect to facets in graphics, and he used `/` and `*` instead for nested and crossed expressions. A crossing structure displays all unique and possible combinations of crossed variables, including those not found in the data, in a graphical layout. However, tidy temporal data finds only the combinations that occur in the data and intentionally respects the structure of incomplete combinations.nclude future time path and distributions generating prediction intervals.
